---
title: "Chapter 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###### ggplot2: Chapter 3 ######

```{r}
rm(list = ls())
setwd("~/Documents/Ggplot2/")
require(ggplot2)
require(dplyr)
require(nlme)
require(maps)
require(USAboundaries)
```

We want each layer of the plot to have a purpose. There are three general purposes for a layer.

1. To display raw data points. This can give an understanding of structure and outliers.
2. To display a statistical summary. This can be model predictions, or mean, median, standard deviation type statistics.
3. Add metadata. This can provide context and references or be used to highlight particulary important features.

Generally we build our plot in order from step 1 to step 3.

_Exercises_

1. What geoms would you use to draw each of the following named plots?  
geom_point() or geom_jitter(), geom_line(), geom_histogram(), geom_bar(), geom_bar() with coord_polar()

2. What's the difference between geom_path() and geom_polygon()? What's the difference between geom_path() and geom_line()?  
geom_path() is not filled in whereas geom_polygon() is. geom_path() can draw a line in any direction whereas geom_line() draws only from left to right.

3. What low-level geoms are used to draw geom_smooth()? What about geom_boxplot() and geom_violin()?  
At a low level, geom_smooth() is just a geom_line() but finds it's slope at a given point using local averages. geom_boxplot() is just a version of geom_raster(). geom_violin() is geom_area().

```{r}
ggplot(Oxboys, aes(age, height, group=Subject)) +
  geom_point() +
  geom_line()

ggplot(Oxboys, aes(Occasion, height)) +
  geom_boxplot() +
  geom_line(aes(group = Subject), color = 'red', alpha = 0.2)
```

```{r}
ggplot(mpg, aes(class, fill=drv)) +
  geom_bar()
```

_Exercises_

1. Draw a boxplot of hwy for each value of cyl, without turning cyl into a factor. What extra aesthetic do you need to set?
```{r}
ggplot(mpg, aes(cyl, hwy, group=cyl)) +
  geom_boxplot()
```

2. Modify the following plot so that you get one boxplot per integer value value of displ. ggplot(mpg, aes(displ, cty)) + geom_boxplot()
```{r}
ggplot(mpg, aes(displ, cty, group=round(mpg$displ))) + geom_boxplot()
```

3. When illustrating the difference between mapping continuous and discrete colours to a line, the discrete example needed aes(group = 1). Why? What happens if that is omitted? What’s the difference between aes(group = 1) and aes(group = 2)? Why?
```{r}
df <- data.frame(x = 1:3, y = 1:3, colour = c(1,3,5))

ggplot(df, aes(x, y, color = factor(colour))) +
  geom_line(aes(group=1)) +
  geom_point(size = 5)
```

4. How many bars are in each of the following plots?
```{r}
ggplot(mpg, aes(drv)) +
  geom_bar()
ggplot(mpg, aes(drv, fill = hwy, group = hwy)) +
  geom_bar()
library(dplyr)
mpg2 <- mpg %>% arrange(hwy) %>% mutate(id = seq_along(hwy))
ggplot(mpg2, aes(drv, fill = hwy, group = id)) +
geom_bar()
```
5. Install the babynames package. It contains data about the popularity of babynames in the US Run the following code and fix the resulting graph. Why does this graph make me unhappy?
```{r}
library(babynames)
hadley <- dplyr::filter(babynames, name == "Hadley")
ggplot(hadley, aes(year, n, color=sex)) +
geom_line()
```

While there is no distinct 3D plotting in ggplot, we can use 2D plots to simulate 3D surfaces.

```{r}
ggplot(faithfuld, aes(eruptions, waiting)) +
  geom_contour(aes(z=density, color=..level..))

ggplot(faithfuld, aes(eruptions, waiting)) +
  geom_raster(aes(fill=density))

ggplot(sample_n(faithfuld, 500), aes(eruptions, waiting)) +
  geom_point(aes(size = density), alpha = 1/3) +
  scale_size_area()
```

We can also visualize spatial data using maps.

```{r}
wi_counties = map_data("county", "wisconsin") %>%
  select(lon=long, lat, group, id=subregion)

head(wi_counties)

ggplot(wi_counties, aes(lon, lat)) +
  geom_polygon(aes(group=group), fill=NA, color="grey") +
  coord_quickmap()
```

```{r}
wi_counties <- map_data("county", "wisconsin") %>%
  select(lon = long, lat, group, id = subregion)

wi_cities <- maps::us.cities %>%
  tbl_df() %>%
  filter(country.etc == "WI") %>%
  select(-country.etc, lon = long) %>%
  arrange(desc(pop))
wi_cities

ggplot(wi_cities, aes(lon, lat)) +
  geom_polygon(aes(group=group), wi_counties, fill=NA, color="grey") +
  geom_point(aes(size = pop)) +
  scale_size_area() +
  coord_quickmap()
```

```{r}
# area metadata

wi_census = midwest %>% 
  tbl_df() %>% 
  filter(state == 'WI') %>%
  mutate(county = tolower(county)) %>% 
  select(county, area, poptotal, percwhite, percblack)

census_counties = left_join(wi_census, wi_counties, by = c("county" = "id"))

ggplot(census_counties, aes(lon, lat, group = county)) +
  geom_polygon(aes(fill = poptotal)) +
  coord_quickmap()
```

```{r}
y = c(18,11,16)
df = data.frame(x=1:3, y=y, se = c(1.2,.5,1))

p = ggplot(df, aes(x,y,ymin=y-se, ymax=y+se))
p+geom_crossbar()
p+geom_pointrange()
p+geom_smooth(stat="identity")
```

```{r}
ggplot(midwest, aes(percwhite, percbelowpoverty)) +
  geom_point(aes(size=poptotal/1e6)) +
  scale_size_area("population\n(millions", breaks=c(.5,1,2,4)) +
  geom_smooth(aes(weight=poptotal), method=lm, size=1)
```

```{r}
ggplot(midwest, aes(percbelowpoverty)) +
  geom_histogram(aes(weight=poptotal))
```

```{r}
head(diamonds)
```

Visualizing larger data sets like the diamonds set for instance can be tricky.

```{r}
ggplot(diamonds, aes(depth)) + 
  geom_histogram(binwidth=0.1) + #Remember to set a custom binwidth. Each data set requires a hand picked binwidth.
  xlim(55,70)
```

Comparing groups:

```{r}
ggplot(diamonds, aes(depth)) +
  geom_histogram(binwidth=0.1) +
  xlim(55,70) +
  facet_wrap(~cut)

ggplot(diamonds, aes(depth)) +
  geom_freqpoly(aes(color=cut), binwidth=0.1) +
  xlim(55,70) 

ggplot(diamonds, aes(depth)) +  #This one is weird but does a good job showing the breakdown at a given point on the x axis
  geom_histogram(aes(color=cut), binwidth=0.1, position="fill") +
  xlim(55,70) 

ggplot(diamonds, aes(depth, fill=cut, color=cut)) +
  geom_density(alpha=0.2) +
  xlim(55,70)
```

_Exercises_

1. What binwidth tells you the most interesting story about the distribution of carat?
```{r}
ggplot(diamonds, aes(carat)) +
  geom_histogram(binwidth=.01)
```

I think a very narrow binwidth tells the most interesting story. You see almost an asymptote at the many of the nicely numbered carat value and a tail off following. This shows that it is very rare to cut a diamond to be just below a nice number.

2. Draw a histogram of price. What interesting patterns do you see?
```{r}
ggplot(diamonds, aes(price)) +
  geom_histogram(binwidth=500)
```

In general there is a very gradual tail off. About the same number of diamonds exist in the 10000 price range as in the 17000 price range. Looks like exponential growth as it goes towards 0 dollars.

3. How does the distribution of price vary with clarity?
```{r}
ggplot(diamonds, aes(clarity, price)) +
  geom_violin()
```

This is very odd. It seems like as the clarity gets greater, the average price lowers. All clarity levels have outliers that stretch beyond 15000 dollars.

4. Overlay a frequency polygon and density plot of depth. What computed variable do you need to map to y to make the two plots comparable?
```{r}
# ggplot(diamonds, aes(depth)) +
#   geom_freqpoly() +
#   geom_density(aes(density*count))
```

```{r}
# dealing with overplotting

# smaller datasets
df = data.frame(x=rnorm(2000), y=rnorm(2000))
ggplot(df, aes(x,y)) +
  geom_point() #overplotted
ggplot(df, aes(x,y)) +
  geom_point(shape=1) #better
ggplot(df, aes(x,y)) +
  geom_point(shape=".") #good on small windows, terrible on large screens


# larger datasets
ggplot(df, aes(x,y)) +
  geom_point(alpha=.3)
```

