---
title: "Chapter5"
author: "Bennett Majerowski"
date: "2/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
```

The goal of ggplot is the provide a framework to make plots through iteration; to be able to build a complex plot layer on top of layer.

Remember from earlier, every plot has:
  1. Data - the underlying data set
  2. Aesthetics - the space that we are mapping the data
  3. Geom - the geometric representation of a datapoint
  
Within a geom, there are actually 5 components:
  1. Mapping - the set of aesthetic mappings specific to this layer. Will be combined with the aesthetic set from the plot defaults.
  2. Data - a layer dataset to override the  plot dataset
  3. Geom - the name of the geometric object to draw each observation
  4. Stat - the statistical transformation to use
  5. Position - the method to adjust overlapping objects
  
__Exercises__

1. Use the result of the code below recreate the pictured plot.

```{r}
class = mpg %>% 
  group_by(class) %>% 
  summarize(n=n(), hwy=mean(hwy))

ggplot(mpg, aes(class, hwy)) +
  geom_jitter(width=0.1) +
  geom_point(data=class, aes(class, hwy), color="red", size=4) +
  geom_text(data=class, aes(class, 10, label=paste("n = ",n, sep="")), size=3)
```

Aestheics describe how a variable is mapped to visual properties. *Note, never refer to a variable using a $ within aes().

Plot aes() vs layer aes():
  1. If you only have a layer, then the aes() info can be put either in the plot or the layer and you will get the same chart.
  2. Conversely, if you have multiple layers, the plot aes() will apply to all layers while the layer aes() will only apply to that layer.
  
Setting vs Mapping:
  1. aes(color=cut) will map the values of the variable "cut" to different colors.
  2. color=cut will set the color for that layer to a constant
  
__Exercises__
1. Simplify the following plot specifications:
```{r}
# bad version
ggplot(mpg) +
  geom_point(aes(mpg$displ, mpg$hwy))

# better version
ggplot(mpg, aes(displ, hwy)) +
  geom_point()
```

```{r}
# bad version
ggplot() +
  geom_point(mapping = aes(y = hwy, x = cty), data = mpg) +
  geom_smooth(data = mpg, mapping = aes(cty, hwy))

# better version
ggplot(mpg, aes(cty, hwy)) +
  geom_point() +
  geom_smooth()
```

```{r}
# bad version
ggplot(diamonds, aes(carat, price)) +
  geom_point(aes(log(brainwt), log(bodywt)), data = msleep)

# better version
ggplot(msleep, aes(log(brainwt), log(bodywt))) +
  geom_point()
```

2. What does the following code do? Does it work? Does it make sense? Why/why not?
```{r}
ggplot(mpg) +
  geom_point(aes(class, cty)) +
  geom_boxplot(aes(trans, hwy))
```

This code does not make sense because it uses one layer with x axis being car class and the other layer has x axis as transmission type. It is essentially two unrelated plots shoved into the same chart.

3. What happens if you try to use a continuous variable on the x axis in one layer, and a categorical variable in another layer? What happens if you do it in the opposite order?
```{r}
ggplot(mpg) +
  geom_point(aes(class, cty)) +
  geom_boxplot(aes(hwy, cty))
```

You cannot plot a categorical layer above a continuous layer. If you plot a continuous layer above a categorical layer, the plot treats the entire continuous layer as one category.

__Exercises__
1. For each of the plots below, identify the geom used to draw it.
  a. violin
  b. point
  c. hex
  d. point
  e. ribbon
  f. path
  
2. For each of the following problems, suggest a useful geom:
  a. Display how a variable has changed over time - line
  b. Show the detailed distribution of a single variable - histogram
  c. Focus attention on the overall trend in a large dataset - smooth
  d. Draw a map - map or polygon
  e. Label outlying points - text
  
Using the stat parameter, we can recreate one of the above charts without needing two data sets.

```{r}
ggplot(mpg, aes(class, hwy)) +
  geom_jitter(width = 0.1) +
  geom_point(stat = "summary", fun.y = "mean", color="red", size=4)
```

__Exercises__

1. The code below creates a similar dataset to stat_smooth(). Use the appropriate geoms to mimic the default geom_smooth().

```{r}
mod <- loess(hwy ~ displ, data = mpg)
smoothed <- data.frame(displ = seq(1.6, 7, length = 50)) 
pred <- predict(mod, newdata = smoothed, se = TRUE) 
smoothed$hwy <- pred$fit
smoothed$hwy_lwr <- pred$fit - 1.96 * pred$se.fit 
smoothed$hwy_upr <- pred$fit + 1.96 * pred$se.fit

ggplot(smoothed, aes(displ)) +
  geom_line(aes(y=hwy)) +
  geom_ribbon(aes(displ, ymin=hwy_lwr, ymax=hwy_upr), alpha=0.3, fill="blue")
```

2. What stats were used to create the following plots?
  a. stat_sum()
  b. stat_qq()
  c. stat_summary()
  
3. Read the docs for stat_sum() then use geom_count() to create a plot that shows the proportion of cars that have each combination of drv and trans.

```{r}
ggplot(mpg, aes(drv, trans)) +
  geom_count()
```

Position adjustments can tweak the position of elements within a layer. For a bar type chart, we have stack (overlapping bars on top of each other), fill (overlapping bars scaled so the top is at 1), dodge (shift overlapping bars side by side)

__Exercises__

1. When might you use position_nudge()?

You can use this when you want a point label to not overlap the point.

2. Many position adjustments can only be used with a few geoms. For example, you can't stack boxplots or error bars. Why not? What properties must a geom possess in order to be stackable? What properties must it possess to be dodgeable?

A geom that that does not start from y=0 cannot be stackable. A geom is not dodgeable if the x-value is very important. Dodging will alter the x values.

3. Why might you use geom_jitter() instead of geom_count()? What are the advantages and disadvantages of each technique?

If there are a lot of overlapping points, geom_jitter allows a simpler way to spread the points out.

4. When might you use a stacked area plot? What are the advantages and disadvantages compared to a line plot?

Stacking area plots can give the user an idea of the aggregate value with all of the groups together as well as a breakdown per group.

