---
title: "Chapter 6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
require(tidyverse)
```

Scales control the mapping from data to aesthetics. ggplot gives default scales but if you want to override them you can.

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color=class)) +
  scale_x_continuous("A really awesome x axis label") +
  scale_y_continuous("An amazingly great y axis label")
```

Scales are x and y as well as color, size and shape. Naming conventions go as follows:  
1. scale_
2. the name of the aes (color, shape, x, etc.)
3. the name of the scale (continuous, discrete, brewer)

__Exercises__  
1. What happens if you pair a discrete variable to a continuous scale? What about the opposite?
```{r}
ggplot(mpg, aes(trans, hwy)) +
  scale_x_continuous("cylinders") +
  geom_point()

ggplot(mpg, aes(trans, hwy)) +
  scale_y_discrete("cylinders") +
  geom_point()
```

If the values for the discrete can be numeric, plotting on a continuous scale will just plot the numeric values of the variables. If the variables are not numeric, then you cannot plot them on a continuous scale.  
If you have a continuous variable plotted on a discrete scale, then each unique value for that variable will be considered its own class.

2. Simplify the following plot specifications to make them easier to understand.
```{r}
ggplot(mpg, aes(displ)) +
  scale_y_continuous("Highway mpg") +
  scale_x_continuous() +
  geom_point(aes(y = hwy))

ggplot(mpg, aes(displ, hwy)) +
  scale_y_continuous("Highway mpg") +
  geom_point()
```

```{r}
ggplot(mpg, aes(y=displ, x=class)) +
  scale_y_continuous("Displacement (l)") +
  scale_x_continuous("Car type") +
  scale_x_continuous("Type of car") +
  scale_color_discrete() +
  geom_point(aes(color=drv)) +
  scale_color_discrete("Drive\ntrain")

ggplot(mpg, aes(class, displ)) +
  scale_x_discrete("Type of car") +
  scale_y_continuous("Displacement (l)") +
  scale_color_discrete("Drive\ntrain") +
  geom_point(aes(color=drv))
```

A guide is the set of labels associated with a certain scale. For instance, a x axis scale will have an axis label, a set of ticks or gridlines, and a set of tick labels. A color scale will have a legend title, key set, and key labels.

In any of the scale functions, the three parameters are the same:
  1. Name. This is the axis or legend title. Just write it in a text string.
  2. Breaks. This determines the values that appear as tick marks on the axes or keys on legends.
  3. Labels. This changes the words that accompany tick marks or keys.
  
Shortcut - just use xlab() to alter just the label of the x axis instead of having to use the scale function.

__Exercises__

1. Recreate the following graphic.
```{r}
ggplot(mpg, aes(displ, hwy)) +
  ylab(expression(paste("Highway (", frac(miles,gallons), ")"))) +
  scale_x_continuous(labels = scales::unit_format("L", 2)) +
  geom_point()
```