---
title: "Chapter 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###### ggplot2: Chapter 3 ######

```{r}
rm(list = ls())
setwd("~/Documents/Ggplot2/")
require(ggplot2)
require(dplyr)
require(nlme)
```

We want each layer of the plot to have a purpose. There are three general purposes for a layer.

1. To display raw data points. This can give an understanding of structure and outliers.
2. To display a statistical summary. This can be model predictions, or mean, median, standard deviation type statistics.
3. Add metadata. This can provide context and references or be used to highlight particulary important features.

Generally we build our plot in order from step 1 to step 3.

_Exercises_

1. What geoms would you use to draw each of the following named plots?  
geom_point() or geom_jitter(), geom_line(), geom_histogram(), geom_bar(), geom_bar() with coord_polar()

2. What's the difference between geom_path() and geom_polygon()? What's the difference between geom_path() and geom_line()?  
geom_path() is not filled in whereas geom_polygon() is. geom_path() can draw a line in any direction whereas geom_line() draws only from left to right.

3. What low-level geoms are used to draw geom_smooth()? What about geom_boxplot() and geom_violin()?  
At a low level, geom_smooth() is just a geom_line() but finds it's slope at a given point using local averages. geom_boxplot() is just a version of geom_raster(). geom_violin() is geom_area().

```{r}
ggplot(Oxboys, aes(age, height, group=Subject)) +
  geom_point() +
  geom_line()

ggplot(Oxboys, aes(Occasion, height)) +
  geom_boxplot() +
  geom_line(aes(group = Subject), color = 'red', alpha = 0.2)
```

```{r}
ggplot(mpg, aes(class, fill=drv)) +
  geom_bar()
```

_Exercises_

1. Draw a boxplot of hwy for each value of cyl, without turning cyl into a factor. What extra aesthetic do you need to set?
```{r}
ggplot(mpg, aes(cyl, hwy, group=cyl)) +
  geom_boxplot()
```

2. Modify the following plot so that you get one boxplot per integer value value of displ. ggplot(mpg, aes(displ, cty)) + geom_boxplot()
```{r}
ggplot(mpg, aes(displ, cty, group=round(mpg$displ))) + geom_boxplot()
```

3. When illustrating the difference between mapping continuous and discrete colours to a line, the discrete example needed aes(group = 1). Why? What happens if that is omitted? What’s the difference between aes(group = 1) and aes(group = 2)? Why?
```{r}
df <- data.frame(x = 1:3, y = 1:3, colour = c(1,3,5))

ggplot(df, aes(x, y, color = factor(colour))) +
  geom_line(aes(group=1)) +
  geom_point(size = 5)
```

4. How many bars are in each of the following plots?
```{r}
ggplot(mpg, aes(drv)) +
  geom_bar()
ggplot(mpg, aes(drv, fill = hwy, group = hwy)) +
  geom_bar()
library(dplyr)
mpg2 <- mpg %>% arrange(hwy) %>% mutate(id = seq_along(hwy))
ggplot(mpg2, aes(drv, fill = hwy, group = id)) +
geom_bar()
```
5. Install the babynames package. It contains data about the popularity of babynames in the US Run the following code and fix the resulting graph. Why does this graph make me unhappy?
```{r}
library(babynames)
hadley <- dplyr::filter(babynames, name == "Hadley")
ggplot(hadley, aes(year, n, color=sex)) +
geom_line()
```